---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## 1. Loading and preprocessing the data 
### Pre-requisite
Data file *activity.csv* (uncompressed version of the *[activity.zip](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)*) has to be preset in the working folder. 

### Loading 
Data is loaded into *activity_data_raw* data frame.
```{r}
activity_data_raw <- read.csv('activity.csv')
head(activity_data_raw)
nrow(activity_data_raw)
```
### Preprocessing 
For later use *activity_data_without_na* data frame is created by removing rows with NA values (can be verfied by comparing the output of *nrow* function)
```{r}
activity_data_without_na <- na.omit(activity_data_raw)
head(activity_data_without_na)
nrow(activity_data_without_na)
```

## 2. What is mean total number of steps taken per day?
The total number of steps for each day is calculated by creating data frame *activity_per_day* ,by using *aggregate* function to sum up total steps taken each day 
```{r}
activity_per_day <- aggregate(steps ~ date, data = activity_data_raw, FUN = sum)
head(activity_per_day)
```
The histogram showing total number of steps each day has been created using *hist* plotting function
```{r}
hist(activity_per_day$steps, xlab = "The total num. of steps per day", main = NULL, col = "green")
```

The mean is calculated using R's *mean* function
```{r}
mean(activity_per_day$steps)
```
The median is calculated using R's *median* function
```{r}
median(activity_per_day$steps)
```

## 3. What is the average daily activity pattern?



## 4. Imputing missing values



## 5. Are there differences in activity patterns between weekdays and weekends?
