---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## 1. Loading and preprocessing the data 
### Pre-requisite
Data file *activity.csv* (uncompressed version of the *[activity.zip](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)*) has to be preset in the working folder. 

Library dplyr has been loaded


### Loading 
Data is loaded into *activity* data frame.
```{r}
library(dplyr)
library(ggplot2)
library(lattice)
activity <- read.csv('activity.csv')
head(activity)
nrow(activity)
```

## 2. What is mean total number of steps taken per day?
The total number of steps for each day is calculated by creating data frame *activity_per_day* ,by using *aggregate* function to sum up total steps taken each day 
```{r}
activity_per_day <- aggregate(steps ~ date, data = activity, FUN = sum, na.rm = TRUE)
head(activity_per_day)
```
The histogram showing total number of steps each day has been created using *hist* plotting function
```{r}
hist(activity_per_day$steps, main = "The total number of steps per day", xlab = "Number of steps", col = "green")
```

The mean is calculated using R's *mean* function
```{r}
mean(activity_per_day$steps)
```
The median is calculated using R's *median* function
```{r}
median(activity_per_day$steps)
```

## 3. What is the average daily activity pattern?

To show the average daily activity pattern, we make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
activity_per_interval <- aggregate(steps ~ interval, data = activity, FUN = mean)

plot(x = activity_per_interval$interval,
     y = activity_per_interval$steps, 
     type = "l",
     main = "Average Daily Activity Pattern",
     xlab = "Interval (mins)",
     ylab = "Number of Steps")
```

Next, we find out which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_avg_daily_activity = activity_per_interval[which.max(activity_per_interval$steps),]
max_avg_daily_activity$interval
```

We have found that interval 835 has the highest average number of steps

## 4. Imputing missing values

The total number of missing values in the dataset (i.e. the total number of rows with NAs) is found to be 2304, by executing the following command:
```{r}
nrow(activity[is.na(activity$steps),])
```

The strategy to replace the missing values is to replace them with the mean number of steps taken from the corresponding interval. 
```{r}
activity_filled <- activity
avg_activity_by_interval <- aggregate(steps ~ interval, data = activity_filled, FUN = mean, rm.na = TRUE)
activity_filled <- merge(activity_filled, avg_activity_by_interval, by="interval", suffixes=c("",".by_interval"))
nas <- is.na(activity_filled$steps)
activity_filled$steps[nas] <- activity_filled$steps.by_interval[nas]
activity_filled <- activity_filled[, c(1:3)]
steps_per_day <- aggregate(steps ~ date, data = activity_filled, FUN = sum)

hist(steps_per_day$steps,breaks=20,labels=unique(steps_per_day$steps[order(steps_per_day$steps)]),main="Histogram of steps by day after Imputing",xlab="Steps")
mean(steps_per_day$steps)
median(steps_per_day$steps)
```
There does not appear to be the significant difference in te mean and median values. However, the histogram is showing that maxium frequency and distribution of data is different.

## 5. Are there differences in activity patterns between weekdays and weekends?
```{r}
# Get day from date
activity_filled$day <- weekdays(as.Date(activity_filled$date))
activity_filled$daytype <- factor(activity_filled$day)
levels(activity_filled$daytype) <- list(
  weekday = c("Monday","Tuesday","Wednesday","Thursday","Friday"),
  weekend = c("Saturday","Sunday")
)

avg_steps_weekdays_vs_weekends <- aggregate(steps ~ daytype + interval, activity_filled, mean)

xyplot(
  type = "l",
  data = avg_steps_weekdays_vs_weekends,
  steps ~ interval | daytype,
  xlab = "Interval",
  ylab = "Number of steps",
  layout = c(1,2)
)

```
There are differences in the activity pattern between weekday and weekend.